Recurrent event analysis in the presence of real-time high frequency data via random subsampling

### 2019-12-02 write-up/fda-recurrent.tex
Making all paragraphs without jumps
updating the tex file

\documentclass[12pt]{amsart}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows}
\usepackage{amsfonts,amssymb,amsthm, txfonts, pxfonts,amscd}
\usepackage[stretch=10]{microtype}
%\usepackage{hyperref}
\def\struckint{\mathop{%
\def\mathpalette##1##2{\mathchoice{##1\displaystyle##2}%
 {##1\textstyle##2}{##1\scriptstyle##2}{##1\scriptscriptstyle##2}}%
\mathpalette
{\vbox\bgroup\baselineskip0pt\lineskiplimit-1000pt\lineskip-1000pt
\halign\bgroup\hfill$}
{##$\hfill\cr{\intop}\cr\diagup\cr\egroup\egroup}%
}\limits}
\usepackage{natbib}
\usepackage{color}
\usepackage{booktabs,caption,fixltx2e}
\usepackage[flushleft]{threeparttable}

@@

Here, we assume each individual is observed over five 12-hour days.
We then generate a sequence of event times via a marginal hazard
rate that ignores sensor data.  We assume the marginal rate is
constant and set to 1 per 12/6/1 hour(s). We then sample non-events at
rates of 1 per 12/6/1/0.5 hours.
This sampling strategy yields (non-)event times~$t$ along with
labels~$Y_t \in \{0,1\}$.
We then use the imputation strategy given by
equation~\eqref{eq:impute} to generate the curves~$\{ X(s,t) \}$.
For simplicity, we do not deal with
We assume that~$Y_t = 0$ implies the curve is flat with mean 0.
We assume that~$Y_t = 1$ implies three curves (A) the curve jumps
around 10 minutes prior to the event and remains elevated, (B) the
curve goes down then up and down smoothly, (C)
These curves are given in Figure~\ref{} for three choices of
parameters that indicate varying degrees of departure from the~$Y_t =
0$ case.
The associated~$\beta$ depends on the covariance~$\Sigma_t (s,t)$.
In particular, $\beta = \Sigma_t^{-1} (\mu_{1,t} - \mu_{0,t})$ is
guaranteed by Lemma~\ref{} and Theorem 1 of \cite{Petrovich2018}.
For~$\beta(s)$ non time-varying implies the covariance term

Using~\eqref{eq:}
We define~$X(t,s)$ at the grid of observerations per secondw ith
\[
  \Sigma_t (s, s^\prime) =
  \frac{\sigma^2}{\Gamma (v) 2^{\nu-1}}
  \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)^{\nu}
  K_v \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)
\]
where~$K_v$ is the modified Bessel function of the second kind, and
we choose~$\nu = 5/2$ and $\sigma^2 = 1$.
We choose three versions of~$\rho_t$: (A) constant across time
at $0.5$, (B) constant within each day but linearly decreasing
across days from~$0.7$ to~$0.3$, and (C) constant within each day but
linearly increasing across days from~$0.3$ to~$0.7$.

@@

Theorem~\ref{thm:biaseddown} shows that if we keep~$K_x = K_b = 35$
and therefore have reconstruction error, we can expect a downward bias
in parameter estimates.
Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias
in the parameter estimates.  For the former, however, we can simply
let~$K_x = K_b$ grow as a function of data size; for the latter, we
can see if~$\hat \beta(s)$ is statistically significant for~$s$
near~$\Delta$. If so, we can set~$\Delta$ larger to ensure the
condition holds.

@@

The imputation strategy generalizes the procedure introduced
by~\cite{Petrovich2018} to the longitudinal functional data
setting.  This is only possible due to the connection established with
logistic regression by Lemma~\ref{lemma:logistic}.
We consider a multiple imputation estimate with~$M$ imputations;
the likelihood estimate is given by~$\bar \theta = M^{-1}
\sum_{m=1}^M \hat \theta^{(m)}$ where~$\theta^{(m)}$ is the estimate
from the~$m$th imputated dataset; the covariance matrix can be
estimated by
\[
M^{-1} \sum_{m=1}^M \hat \Xi^{(m)} (\theta) + \frac{1}{M-1}
\sum_{m=1}^M (\hat \theta^{(m)} - \bar \theta) (\hat \theta^{(m)} -
\bar \theta)
\]
where~$\hat \Xi^{(m)}$ is the estimator~\eqref{eq:fisher} for
the~$m$th imputated dataset.

>>>>

\documentclass[11pt]{amsart}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows}
\usepackage{amsfonts,amssymb,amsthm, txfonts, pxfonts,amscd}
\usepackage[stretch=10]{microtype}
%\usepackage{hyperref}
\def\struckint{\mathop{%
\def\mathpalette##1##2{\mathchoice{##1\displaystyle##2}%
 {##1\textstyle##2}{##1\scriptstyle##2}{##1\scriptscriptstyle##2}}%
\mathpalette
{\vbox\bgroup\baselineskip0pt\lineskiplimit-1000pt\lineskip-1000pt
\halign\bgroup\hfill$}
{##$\hfill\cr{\intop}\cr\diagup\cr\egroup\egroup}%
}\limits}
\usepackage{natbib}
\usepackage{color}
\usepackage{booktabs,caption,fixltx2e}
\usepackage[flushleft]{threeparttable}

@@

Here, we assume each individual is observed over five 12-hour days. We then generate a sequence of event times via a marginal hazard rate that ignores sensor data.  We assume the marginal rate is
constant and set to 1 per 12/6/1 hour(s). We then sample non-events at rates of 1 per 12/6/1/0.5 hours. This sampling strategy yields (non-)event times~$t$ along with labels~$Y_t \in \{0,1\}$. We then use the imputation strategy given by equation~\eqref{eq:impute} to generate the curves~$\{ X(s,t) \}$. For simplicity, we do not deal with We assume that~$Y_t = 0$ implies the curve is flat with mean 0. We assume that~$Y_t = 1$ implies three curves (A) the curve jumps around 10 minutes prior to the event and remains elevated, (B) the curve goes down then up and down smoothly, (C) These curves are given in Figure~XX for three choices of parameters that indicate varying degrees of departure from the~$Y_t = 0$ case. The associated~$\beta$ depends on the covariance~$\Sigma_t (s,t)$. In particular, $\beta = \Sigma_t^{-1} (\mu_{1,t} - \mu_{0,t})$ is guaranteed by Lemma~XX and Theorem XX of~\cite{Petrovich2018}. For~$\beta(s)$ non time-varying implies the covariance term. We define~$X(t,s)$ at the grid of observerations per second with
\[
  \Sigma_t (s, s^\prime) =
  \frac{\sigma^2}{\Gamma (v) 2^{\nu-1}}
  \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)^{\nu}
  K_v \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)
\]
where~$K_v$ is the modified Bessel function of the second kind, and
we choose~$\nu = 5/2$ and $\sigma^2 = 1$.
We choose three versions of~$\rho_t$: (A) constant across time
at $0.5$, (B) constant within each day but linearly decreasing
across days from~$0.7$ to~$0.3$, and (C) constant within each day but
linearly increasing across days from~$0.3$ to~$0.7$.

@@

Theorem~XX shows that if we keep~$K_x = K_b = 35$ and therefore have reconstruction error, we can expect a downward bias in parameter estimates. Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias in the parameter estimates.  For the former, however, we can simply let~$K_x = K_b$ grow as a function of data size; for the latter, we can see if~$\hat \beta(s)$ is statistically significant for~$s$ near~$\Delta$. If so, we can set~$\Delta$ larger to ensure the condition holds.

@@

The imputation strategy generalizes the procedure introduced by~\cite{Petrovich2018} to the longitudinal functional data setting.  This is only possible due to the connection established with logistic regression by Lemma~\ref{lemma:logistic}. We could consider a multiple imputation estimate with~$M$ imputations; the likelihood estimate is given by~$\bar \theta = M^{-1} \sum_{m=1}^M \hat \theta^{(m)}$ where~$\theta^{(m)}$ is the estimate from the~$m$th imputated dataset; the covariance matrix can be estimated by
\[
M^{-1} \sum_{m=1}^M \hat \Xi^{(m)} (\theta) + \frac{1}{M-1}
\sum_{m=1}^M (\hat \theta^{(m)} - \bar \theta) (\hat \theta^{(m)} -
\bar \theta)
\]
where~$\hat \Xi^{(m)}$ is the estimator~\eqref{eq:fisher} for
the~$m$th imputated dataset.

\subsection{Bootstrap and multiple imputation under uncongenial }

Positives: (1) our imputation model is ``approximately'' congenial; (2) our imputatoin model is simple and easy to use.
Negatives: (1) clearly not congenial under many events and/or high sampling rates. (2) difficult when we overlap due to over-sampling.

Congeniality matters because we want to ensure good coverage
Q: Can we derive a variance that provides coverage
We know our imputation models are uncongenial albeit simple to implement.  Impact should be small if events are rare as impute only in ranges where 1,0 labels give you everything.  To be careful, we investigate alternatives and assess impact.

Simple version compared to the von Hippel's boot MI proposal.
\begin{itemize}
\item Bootstrap $B$ times
\item For $b = 1, \ldots, B$, impute $M$ times
\item Let~$\hat \theta_b = M^{-1} \sum_m \hat \theta_{b,m}$
\item Compute $\text{Var}_{BootMI} = (B-1)^{-1} \sum_{b=1}^B (\hat \theta_b - \hat \theta_{BM})^2$ where $\hat \theta_{BM} = B^{-1} \sum_{b=1}^B \hat \theta_b$.
\end{itemize}
